---
title: "Activity_prediction"
author: "Víctor Zambrano"
date: "3 de abril de 2016"
output: html_document
---

## Summary
The following document is the result of the anaylis made on the data made public
by a research team on the Pontifical Catholic University of Rio de Janeiro, Brazil.
The purpose of the analysis was to device a predicting model based on machine learning,
and the chosen algorithm to do so was 
[linear discriminant analysis](https://en.wikipedia.org/wiki/Linear_discriminant_analysis)


## Data Loading and splitting
First we load the project data:

```{r}
full_data <- read.csv('./data/pml-training.csv')
final_test <- read.csv('./data/pml-testing.csv')
```

Now we split the data between training and testing for Cross-validation purposes

```{r}
library(caret)
inTrain <- createDataPartition(y=full_data$classe, p=0.75, list=FALSE)
training <- full_data[inTrain,]
testing <- full_data[-inTrain,]
```

## Data Pre-processing
The working data has 160 variables and it would not be wise to build the model
with all those variables, so now we take a look at the data to filter it. 

```{r, results='hide'}
summary(training)
```

For legibility purposes, the output of the last command was hidden. It showed
the summary of 160 variables, some of which took the `NA` value a lot of
times, therefore we'll filter those out.

```{r}
number_of_na <- as.data.frame(sapply(training, function(x) {sum(is.na(x))}))
colnames(number_of_na)[1] <- 'Number_of_NAs'
rows_with_few_NAs <- rownames(number_of_na)[number_of_na$Number_of_NAs<14000]
train_subset <- training[,rows_with_few_NAs]
```

That last preprocessing step has to be repeated in the `testing` Data Frame

```{r}
test_subset <- testing[,rows_with_few_NAs]
```

Now we take a look at the filtered data

```{r, results='hide'}
summary(train_subset)
```

Again... the output of the last command was hidden. It showed the summary of
93 variables (67 were filtered out in the last preprocessing step). Many of the
remaining variables are factors that take the `''` value several times. We'll
filter those out.

```{r}
number_of_empty <- as.data.frame(sapply(train_subset, function(x) {sum(x=='')}))
colnames(number_of_empty)[1] <- 'Number_of_empty'
rows_with_few_empty <- rownames(number_of_empty)[number_of_empty$Number_of_empty<14000]
train_subset <- train_subset[,rows_with_few_empty]
test_subset <- test_subset[,rows_with_few_empty]
```

## Model Training
For this particular case (an analysis done in MY laptop), computing power is an issue.
So the algorithm of choice will be linear discriminant analysis, which is a computationally
convenient algorithm. Moreover, the data will be further preprocessed by performing
a principal component analysis.


```{r}
regreLda2 <- train(classe ~ ., method='lda', preProcess='pca', data = train_subset)
regreLda2
```

## Accuracy estimate
The previous summary shows an optimistic estimate on the model's accuracy. A more realistic
estimate is given by the `confusionMatrix` function.

```{r}
confusionMatrix(test_subset$classe, predict(regreLda2, test_subset))
```

Finally, the Confusion Matrix allows us to estimate the model's accuracy on 82%. The predictions
for the testing set are as follows:

```{r}
predict(regreLda2, final_test)
```